Efficient Handling of String-Number Conversion

### 2019-11-23 draft.tex
sec 5

	\end{tikzpicture} }
\paragraph{Flat automata} 
A finite state automaton  
$A = (Q,T,\Sigma,\st{0}{0},\st{m}{0})$ is said to be \emph{flat} if it satisfies the following structural constraints (the figure above):
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a straight path of $m-1$ transitions $(\st{i}{0},a_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. 
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, a_{i}^{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that the case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
	\item Each character in $\Sigma$ appears on at most one transition. 
\end{enumerate}

The crucial feature of flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. Such encoding into linear formula results in efficient algorithms and decision procedure for the class of flat automata and so we avoid dealing with costly standard automata operations (e.g., checking  the non-emptiness of the intersection of several regular languages
 is known to be {\sc Pspace}-complete while it is in {\sc NP} for the class of flat automata).
This is possible thanks to the flatness, which causes that 
every word $w\in L(A)$ is uniquely determined by its Parikh image $\parikhof{w}$. 
More precisely, the Parikh image of a word $w\in L(A)$ can be seen as an encoding of $w$ and can be uniquely decoded:

\begin{lemma}\label{lemma:decoding}
For an flat FA $A$, there is a function $\decode A$ such that for each $w\in L(A)$, $\decode A(\parikhof w) = w$. 
\end{lemma}
%
\begin{proof}
First observe that the $\parikhof w$ value of any variable appearing within the a cycle of $A$ is equal to the number of repetitions of that cycle in the accepting run. This is an immediate consequence of the fact that 
 every variable appears on at most one transition. Thus, the accepting run on $w$ (and so $w$ itself) can be reconstructed from $\parikhof w$.  More concretely, the function $\decode A$ can be implemented as follows. 
Given $\pim:\#\Sigma\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m}$, 
$\decode A(\pim)$ is constructed as the word $w_0 a_{0} w_1 \cdots a_{m-1} w_{m}$ where for each $i\in [0,m]$,
$w_i = (a_i^0 \cdots a_i^{l_{i}-1})^{\pvarsof {a_i^0}}$ if $l_i >0$ and $w_i = \epsilon$ if $l_i = 0$. 
\end{proof}
%
For example, in the automaton given at the beginning of this section, 
from $\parikhwof x {a_0} = \parikhwof x {a_1} = \cdots = \parikhwof x {a_{m-1}}=1$, $\parikhwof x {a^0_1}=\parikhwof x {a^1_1}=\parikhwof x {a^2_1}=2$ and $\parikhwof x {a^i_j}=0$ otherwise,  we derive that $x=a_0(a^0_1a^1_1a^2_1)^2a_1\cdots a_{m-1}$.

@@

We say that a mapping $I_e:\cvar\cup\#\cvar \rightarrow \nat$ is a \emph{word encoding} a word $w$  (or a $\pa$-encoding of $w$) if $w$ is an instantiation of some parametric word $x \in L(\pa)$ whose Parikh image and interpretation of character variables are defined by $I_e$.  
%
Conversely, $w$ is a \emph{$\pa$-decoding} of $I_e$.
%
We use $\encode \pa(w)$ below to denote all $\pa$-encodings of a word $w$,
and $\decode \pa (I_e)$ to denote all $\pa$-decodings of a word encoding $I_e$. Namely,  
% to \emph{word encodings} $e:\cvar\cup\#\cvar \rightarrow \nat$ via the PA $\pa$ and the function $\decode \pa$ does it in the reverse direction.
	$$\encode \pa( w ) = \{I_e \mid  x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x}\}$$
	$$\decode \pa(I_e) =\{ w \mid x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x} \}$$

@@

If $\pa$ is a PFA, then by Lemma~\ref{lemma:decoding}, each word encoding $I_e\in \encode \pa( w )$ can be decoded uniquely to the word $w$, i.e.
$$
w = \decode \pa (\encode \pa (w))
$$
Consequently, as stated by the following corollary of Lemma~\ref{lemma:decoding}, 
Parikh image of parametric words in $L(A)$ paired with character variable interpretations satisfying $\paf$ encode precisely the words in $\semof{\pa}$.
%words in $\semof{\pa}$ are precisely the decodings of the Parikh image of parametric words in $L(A)$ paired with character variables interpretations satisfying $\paf$:

>>>>

	\end{tikzpicture} }
\paragraph{Flat automata} 
A finite state automaton  
$A = (Q,T,\Sigma,\st{0}{0},\st{m}{0})$ is said to be \emph{flat} if it satisfies the following structural constraints (see also the figure above):
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a straight path of $m-1$ transitions $(\st{i}{0},a_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. 
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, a_{i}^{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that the case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
	\item Each character in $\Sigma$ appears on at most one transition. 
\end{enumerate}

The crucial feature of flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. Such encoding into linear formula results in efficient algorithms and decision procedures. For instance, we avoid dealing with costly standard automata operations (e.g., checking  the non-emptiness of the intersection of several regular languages is known to be {\sc Pspace}-complete while it is in {\sc NP} for the class of flat automata).
The encoding is possible due to the flat structure, which has the property that 
``every word $w\in L(A)$ is uniquely determined by its Parikh image $\parikhof{w}$". 
More precisely, the Parikh image of a word $w\in L(A)$ can be seen as an encoding of $w$ and can be uniquely decoded:

\begin{lemma}\label{lemma:decoding}
For an flat FA $A$, there is a function $\decode A$ such that for each $w\in L(A)$, $\decode A(\parikhof w) = w$. 
\end{lemma}
%
%\begin{proof}
Observe that the $\parikhof w$ value of any variable appearing within the a cycle of $A$ is equal to the number of repetitions of that cycle in the accepting run. This is an immediate consequence of the fact that 
 every variable appears on at most one transition. Thus, the accepting run on $w$ (and so $w$ itself) can be reconstructed from $\parikhof w$.

 More concretely, the function $\decode A$ can be implemented as follows. 
Given $\pim:\#\Sigma\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m}$, 
$\decode A(\pim)$ is constructed as the word $w_0 a_{0} w_1 \cdots a_{m-1} w_{m}$ where for each $i\in [0,m]$,
$w_i = (a_i^0 \cdots a_i^{l_{i}-1})^{\pvarsof {a_i^0}}$ if $l_i >0$ and $w_i = \epsilon$ if $l_i = 0$. 
%\end{proof}

For example, in the automaton given at the beginning of this section, 
from $\parikhwof x {a_0} = \parikhwof x {a_1} = \cdots = \parikhwof x {a_{m-1}}=1$, $\parikhwof x {a^0_1}=\parikhwof x {a^1_1}=\parikhwof x {a^2_1}=2$ and $\parikhwof x {a^i_j}=0$ otherwise,  we derive that $x=a_0(a^0_1a^1_1a^2_1)^2a_1\cdots a_{m-1}$.

@@

We say that a mapping $I_e:\cvar\cup\#\cvar \rightarrow \nat$ is a \emph{word encoding} a word $w$  (or a $\pa$-encoding of $w$) if $w$ is an instantiation of some parametric word $x \in L(\pa)$ whose Parikh image and interpretation of character variables are defined by $I_e$.  
%
Conversely, $w$ is a \emph{$\pa$-decoding} of $I_e$.
%
We use $\encode \pa(w)$ below to denote all $\pa$-encodings of a word $w$,
and $\decode \pa (I_e)$ to denote all $\pa$-decodings of a word encoding $I_e$. Namely,  
% to \emph{word encodings} $e:\cvar\cup\#\cvar \rightarrow \nat$ via the PA $\pa$ and the function $\decode \pa$ does it in the reverse direction.
	$$\encode \pa( w ) \defeq \{I_e \mid  x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x}\}$$
	$$\decode \pa(I_e) \defeq\{ w \mid x \in L(\pa) , I(x) =w, I\in\semof \paf, I_e = I \cup \parikhof{x} \}$$

@@

If we consider ${\pa_x}=(A_x,\true)$ from Figure~\ref{fig:sfa}~(a) and let $Y{=}(v_1,v_2,\#v_1,\#v_2)$, we have\\
$\encode {\pa_x}( ``aaa" ) =\{ (Y{\rightarrow}(\semof{a},\semof{\epsilon},3,3), Y{\rightarrow}(\semof{\epsilon},\semof{a},3,3)  \}$ and
$\decode {\pa_x}((Y{\rightarrow}(\semof{a},\semof{\epsilon},3,3))=\{``aaa"\} $.

If $\pa$ is a PFA, then by Lemma~\ref{lemma:decoding}, every word encoding $I_e\in \encode \pa( w )$ can be decoded uniquely to the word $w$, i.e.
$$
\{w\} = \decode \pa (\encode \pa (w))
$$

Similarly, as stated by the following corollary of Lemma~\ref{lemma:decoding}, 
Parikh image of parametric words in $L(A)$ paired with character variable interpretations satisfying $\paf$ encode precisely the words in $\semof{\pa}$.
%words in $\semof{\pa}$ are precisely the decodings of the Parikh image of parametric words in $L(A)$ paired with character variables interpretations satisfying $\paf$:
